SSA based global code motion of identical computations

### 2017-04-16 acmart-pacmpl-template.tex
time

In order to hoist a subset of identical instructions, we partition the list of
all candidates in a way to maximize the total number of hoistings.  By sorting
the list of all the candidates in the increasing order of their depth first
search discovery time stamp \cite{clrs} (DFSIn numbers), we make sure that
candidates closer in the list have their common dominator nearby. In
Figure-\ref{fig:dfsin}, if B3, B4, B5, and B6 have identical computations and for
some reason they cannot be hoisted together at B0, then we partition the set of
hoistable candidates in their DFSIn order. In this case the DFSIn ordering would
be B3, B4, B5, B6 which will allow the instructions in B3, B4 to be hoisted at
B1 and those in B5, B6 to be hoisted at B2.

@@

\subsection{Time complexity of algorithm}
The complexity of code hoisting is linear in number of instructions that could
be hoisted in the program, matching the complexity of \PRE{} on \SSA{} form.  The
analysis phase is based on the Global Value Numbering (\GVN{}), the same analysis
used for \PRE{}, followed by the computation of a partition of identical
expressions to be hoisted in a same location to guarantee safety properties and
program performance, and followed by a simple code generation that adds the
identified instruction in the hoisting point and removes all the now redundant
expressions.

>>>>

In order to hoist a subset of identical instructions, we partition the list of
all candidates in a way to maximize the total number of hoistings.  By sorting
the list of all the candidates in the increasing order of their depth first
search discovery time stamp \cite{clrs} (DFSIn numbers), we make sure that
candidates closer in the list have their common dominator nearby. In
Figure-\ref{fig:dfsin}, if B3, B4, B5, and B6 have identical computations and
for some reason they cannot be hoisted together at B0, then we partition the set
of hoistable candidates in their DFSIn order. In this case the DFSIn ordering
would be B3, B4, B5, B6 which will allow the instructions in B3, B4 to be
hoisted at B1 and those in B5, B6 to be hoisted at B2. Even if (B3, B4) could
not be hoisted, for some reason, to B1, (B5, B6) can still be hoisted -- if
legal and profitable -- to B2.

@@

\subsection{Time complexity of algorithm}
The complexity of code hoisting is linear in number of instructions that are
candidates for code-motion, matching the complexity of \PRE{} on \SSA{} form.
The analyses computed for this pass are Global Value Numbering, Computation of
DJ Graph and MergeSets, Marking Barriers, all are linear in number of
instructions in a function. Liveness analysis expensive but only performed
on-demand for hoistable candidates so it does not affect the overall compile
time by much. Other analyses like Alias Analysis, Memory SSA and Dominator Tree
are already available in the LLVM pass pipeline. Although we recompute GVN, and
Barriers for each iteration of the code-motion, we have not seen signinficant
increase in the compilation times. We have also provided appropriate compiler
flags to expedite code-motion by bailing out with fewer iterations, or skip the
liveness based profitability analysis to aggressively move the code as long as
they are legal (Section~\ref{sec:experimental-results}).

The analysis is followed by a simple code generation that adds the identified
instruction in the destination point and removes all instructions rendered
redundant as a result of code-motion.
